---
title: "1R Mammals Report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(OneR)

# included for confusion matrix
library(caret)
library(e1071)

#used for id3
library(rpart)
library(rpart.plot)
```

**Name:**   Austin Sampson

**eMail:** aws9t5@mst.edu
               
**Course:** CS 5402

**Date:**  02-14-2020

## Concept Description:
Train a system from existing data to classify animals as amphibian, anthropod, bird, fish,
insect, mammal or reptile. Calssification will be done through Naive Classification and ID3
and model will be selected after comparing their respective accuracy and error rates to present 
to the empoloyer.

## Data Collection:
Data has been provided from the client based off the observation of their feild agents.
The training and test data sets provided by the client are tranimal-taxonomy-train.csv and
animal-taxonomy-test.csv respectivly.

## Example Description:

**animal.name**

nominal attribute name of the animal or species
This attribute was removed from the training and test data sets to prevent overfitting

**hair**

Nominal boolean attribute that displays output as:
True
False

**feathers**

Nominal boolean attribute that displays output as:
True
False

**eggs**

Nominal boolean attribute that displays output as:
True
False

**milk**

Nominal boolean attribute that displays output as:
True
False

**airborne**

Nominal boolean attribute that displays output as:
True
False

**aquatic**

Nominal boolean attribute that displays output as:
True
False

**preditor**

Nominal boolean attribute that displays output as:
True
False

**toothed**

Nominal boolean attribute that displays output as:
True
False

**backbone**

Nominal boolean attribute that displays output as:
True
False

**breathes**

Nominal boolean attribute that displays output as:
True
False

**venomous**

Nominal boolean attribute that displays output as:
True
False

**fins**

Nominal boolean attribute that displays output as:
True
False

**legs**

Ratio Lable displaying the number of legs. null value or 0
indicates the absince of legs.
This data was filtered out of the test and training data due
to containing zero values.

**tail**

Nominal boolean attribute that displays output as:
True
False

**domestic**

Nominal boolean attribute that displays output as:
True
False

**catsize**

Nominal boolean attribute that displays output as:
True
False

**gestation**

Interval attribute displays a measure of time it took for the gestation of a species.
this data was removed from the training and test data sets do to missing values.

**type**

Nominal, Main classificationn variable for this data set. Output displayed as:
mammal
fish
arthropod
bird
insect
amphibian
reptile

## Data Import and Wrangling:

Importing test and training data
```{r}
#import main file
train <- read.csv("animal-taxonomy-train.csv")
test <- read.csv("animal-taxonomy-test.csv")

```
We remove the name attribute to prevent the moddel from overfitting by training for name attributes.
We also set the legs and gestation attributes to null do to missing and zero values.
```{R}
#remove the name attribute from the test and training data

train$animal.name <- NULL
test$animal.name <- NULL
train$gestation <- NULL
test$gestation <- NULL
train$legs <- NULL
train$legs <- NULL

```

## Mining and Analytics:

**Naive Bayes Classifiers**

Traing the Confusion Matrix for bayes classification
```{r}
train.nb <- naiveBayes(type ~ ., data=train)
train.nb
```


**ID3**

To train the ID3 Classifier I untailzied the rpart package and their graphing tools
```{r}
#train the id3 model using the rpart library
model.id3 <-rpart(type ~., data = train)
rpart.plot(model.id3)
```


## Evaluation:

**ID3**
Calculating the Confusion matrix for ID3 decision tree classification
Using the model create a prediction for the dataset and compare it to the actual

```{r}
#predict classification using id3 model
pred.id3 <- predict(model.id3, test,type="class")
#print the generated classification decision tree

#Calculated the confusion matrix based off the prediction
pred.id3 <- droplevels(pred.id3)
confusionMatrix(pred.id3,test$type)
```



**Bayes**
Calculating confusion matrix for Bayes
I had re-added the levels that where missing from both tables in order to calculate the confusion
matrix. 
```{r}
m_predictions <- predict(train.nb, test)

#table(m_predictions);
#table(test);


m_predictions <- droplevels(m_predictions)
cfm <- confusionMatrix(m_predictions, test$type)
cfm

```

**Accuracy and Error Rates**

As can be shown above due to both the bayes and ID3 algorithm generating
accuracies of 1, we know both models have an error rate of 0%.


**ID3**

## Results

Since both the generated classification modesl produced a 100% accuracy rating
from their confusion matrixes based off our test data and therefore had error
rates of 0%. We can conclude that both the Naive Bayes and ID3 models are equivalent
in their ability to produce accurate and relaible results for the classification of our data.

Out of the two models examened I would recomend presenting the ID3 model to the client due to
its ease of use and graphical representation. Another reason to choose ID3 would be because
it does not operate under the assumption that all the data attributes are independent Like Naive Bayes.


## References:
https://www.rdocumentation.org/packages/caret/versions/3.45/topics/confusionMatrix
https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/data
http://www.learnbymarketing.com/tutorials/naive-bayes-in-r/
https://www.rdocumentation.org/packages/rpart/versions/4.1-15/topics/rpart
http://www.milbo.org/rpart-plot/prp.pdf



